---
title: "Reproducible Research: Peer Assessment 1"
author: "PanMud"
date: "Octorber 1, 2017"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Loading and preprocessing the data

```{r activity}
suppressPackageStartupMessages(library(dplyr))
# Unzip archive
unzip("activity.zip")

# Read base data into a data frame.
activity <- read.csv("activity.csv")

# Data Dimensions
dim(activity)
# Data Summary
summary(activity)
```
## What is mean total number of steps taken per day?

```{r q1}
suppressPackageStartupMessages(library(ggplot2))
total.steps <- tapply(activity$steps, activity$date, FUN=sum, na.rm=TRUE)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day", ylab="Frequency",
      fill=I("blue"), col=I("black"))
```
mean of the total number of steps taken per day

```{r q1a}
mean(total.steps, na.rm=TRUE)
```
median of the total number of steps taken per day

```{r q1b}
median(total.steps, na.rm=TRUE)
```

## What is the average daily activity pattern?

```{r q2}
suppressPackageStartupMessages(library(ggplot2))
averages <- aggregate(x=list(steps=activity$steps), by=list(interval=activity$interval),
                      FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) + geom_line(color="blue")  + 
       xlab("5-minute interval") + ylab("average number of steps taken")
```    
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r q2a}
averages[which.max(averages$steps),]
```

## Imputing missing values
total number of missing values in the dataset 

```{r q3a}
missing <- is.na(activity$steps)
# How many missing
table(missing)
```

Fill the missing values with mean value for that 5-minute interval.

```{r q3b}
# Replace each missing value with the mean value of its 5-minute interval
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}
cleanData <- activity
cleanData$steps <- mapply(fill.value, cleanData$steps, cleanData$interval)
```


Make a histogram of the total number of steps taken each day and calculate the mean and median total number of steps.

```{r q3c}
total.steps <- tapply(cleanData$steps, cleanData$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day", ylab="Frequency", 
fill=I("blue"), col=I("black"))
```

mean of the total number of steps taken per day in the new dataset

```{r q3d}
mean(total.steps, na.rm=TRUE)
```

median of the total number of steps taken per day in the new dataset

```{r q3e}
median(total.steps, na.rm=TRUE)
```
## Are there differences in activity patterns between weekdays and weekends?

Ientify the the weekend and weekdays in the data
```{r q5a}
  wdaywend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
cleanData$date <- as.Date(cleanData$date)
cleanData$day <- sapply(cleanData$date, FUN=wdaywend)
```
Plot the number of steps based on the weekday or weekend

```{r q5b}
averages <- aggregate(steps ~ interval + day, data=cleanData, mean)
ggplot(averages, aes(interval, steps)) + geom_line(color="blue")  + facet_grid(day ~ .) +
    xlab("5-minute interval") + ylab("Number of steps")
```

Note that there seems to be a variation in the activity patterns between weekdays and weekends.
